Staged Abstract Interpreters

### 2018-11-15 paper/23_bg_stage.tex
paper

\begin{lstlisting}
.. = new DslDriver[Int, Int] {
  def power(b: Rep[Int], x: Int): Rep[Int] =
    if (x == 0) 1 else b * power(b, x-1)
  def snippet(x: Rep[Int]): Rep[Int] = power(x, 5)
}
\end{lstlisting}

>>>>

\begin{lstlisting}
new DslDriver[Int, Int] {
  def power(b: Rep[Int], x: Int): Rep[Int] =
    if (x == 0) 1 else b * power(b, x-1)
  def snippet(x: Rep[Int]): Rep[Int] = power(x, 5)
}
\end{lstlisting}

### 2018-11-15 paper/30_sai.tex
paper

When generating code for application, we can not directly apply the callee.
Instead, we emit code that calls a next-stage function @apply_closures_norep@ 
which non-deterministically applies multiple target closure with the argument
and latest store, and finally returns the joined value and a single store.

@@

\subsection{Staged Fixpoint Iteration.} 
\todo{staged fixed point iteration with memoization}

@@

Now we already obtained a working version of staged abstract interpter that is able to 
specialize an analysis. However, we treat the data structures such as @Map@s as black-boxes,
which means any operations on a @Map@ becomes code in the next stage.
But, as we identified before \todo{where}, the keys of environment are identifiers in the program,
which are completely known statically. This leaves us a chance to further specialize the 
data structures. Particularly, if we are specializing a monovariant analysis, the address space 
is equivalent to the identifiers, then the environment component can be entirely eliminated, 
and the store is a partially specialized map.
\todo{maps to arrays}

>>>>

When generating code for application, we can not directly apply the callee.
Instead, we emit code that calls a next-stage function @apply_closures_norep@.
Function @apply_closures_norep@ does the same job as its unstaged version,
which non-deterministically applies multiple target closures with the argument
and latest store, and finally returns the joined value and a single store.
We provide the definition of @apply_closures_norep@ in the runtime supporting code.

@@

\subsection{Staged Fixpoint Iteration}

Our fixed-point iteration relies two caches @in@ and @out@, but the iteration 
can not be done at the current stage.
Because the @in@ and @out@ are both next-stage values, thus the test of whether 
@in@ and @out@ are equal is a generated expression to the next stage, and we can only 
know the comparison result at the next stage. 
In other words, we do not know how many iterations we need to reach the fixed-point.
To achieve this, we need to stage a
function --- @iter_aux@ is generated as a recursive function of type @Rep[Unit => (Value,Store)]@
and will be invoked at the next stage.

\begin{lstlisting}
def iter(e: Expr, $\rho$: Rep[Env], $\sigma$: Rep[Store]): 
Rep[(Value,Store)] = {
  def iter_aux: Rep[Unit => (Value,Store)] = fun { () =>
    in = out; out = Map[Config, (Value,Store)]()
    cached_ev(e, $\rho$, $\sigma$)
    if (in === out) out((unit(e), $\rho$, $\sigma$)) else iter_aux()
  }
  iter_aux() // generated code that invokes iter_aux()
}
\end{lstlisting}

However, the instrumented evaluation function that uses @in@ cache and updates @out@ cache
can be completely eliminated by staging. Each recursive call to @cached_ev@ will also be specialized
if it is applied on subexpressions of the analyzed program.

\begin{lstlisting}
def cached_ev(e: Expr, $\rho$: Rep[Env], $\sigma$: Rep[Store]): 
Rep[(Value, Store)] = {
  val cfg: Rep[Config] = (unit(e), $\rho$, $\sigma$)
  if (out.contains(cfg)) { out(cfg) }
  else {
    val ans0 = in.getOrElse(cfg, RepLattice[(Value, Store)].bot)
    out = out + (cfg -> ans0)
    val ans1 = evev(cached_ev)(e, $\rho$, $\sigma$)
    out = out + (cfg -> (ans0 $\sqcup$ ans1)); ans1
  }
}
\end{lstlisting}

@@

Now we already obtained an end-to-end staged abstract interpreter that is able to 
specialize an analysis. However, we treat the data structures such as @Map@s as black-boxes,
which means any operations on a @Map@ becomes code in the next stage.
But, as we identified before \todo{where}, the keys of environment are identifiers in the program,
which are completely known statically. This leaves us a chance to further specialize the 
data structures. Particularly, if we are specializing a monovariant analysis, the address space 
is equivalent to the identifiers, then the environment component can be entirely eliminated, 
and the store is a partially specialized map.
\todo{maps to arrays}
